---
title: "DESeq2"
author: "Phi Le"
date: "8/7/2021"
output: pdf_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

```{r}
library(GenomicFeatures)

library(stageR)
library(tidyverse)
library(data.table)
library(tximport)
library(DESeq2)
library(apeglm)
library(ggplot2)
library(ggrepel)
library(EnhancedVolcano)
```



# Loading data from salmon results

```{r}
samples <- read.table("C:/Users/phill/Documents/Studies and Research/Project with Kurt/src_speed_814/test/targets.txt", header=TRUE)
samples <- samples %>% dplyr::rename(sample_id = library) %>% filter(group %in% c("HFD_OX", "NMD_OX")) %>% mutate(group = factor(group), sample_id = factor(sample_id)) %>% dplyr::rename( condition = group)

files <- file.path("C:/Users/phill/Documents/Studies and Research/Project with Kurt/src_speed_814/test/data", samples$sample_id, "quant.sf")
files

freq_sample = table(samples$group) # for counting number of sample size

TxDb <- AnnotationDbi::loadDb("test/new_TxDB.txt")
k <- keys(TxDb, keytype = "TXNAME")
txdf <- AnnotationDbi::select(TxDb, k, "GENEID", "TXNAME")

# txi <- tximport(files, type="salmon", tx2gene=txdf, txOut=TRUE,
                # countsFromAbundance="scaledTPM")

txi = tximport(files, type = "salmon", tx2gene = txdf, countsFromAbundance = "no")

ddsTxi <- DESeqDataSetFromTximport(txi,
                                   colData = samples,
                                   design = ~ condition)
```

# Loading data from 
```{r}
library("tximeta")
samples <- read.table("C:/Users/phill/Documents/Studies and Research/Project with Kurt/src_speed_814/test/targets.txt", header=TRUE)

samples <- samples %>% dplyr::rename(id = library, condition = group) %>% dplyr::mutate(line = str_sub(id,1,7))%>% filter(condition %in% c("ETB_KO_HFD", "HFD_OX")) %>% 
  mutate(condition = factor(condition), line = factor(line)) %>% as_tibble()

files <- file.path("C:/Users/phill/Documents/Studies and Research/Project with Kurt/src_speed_814/test/data", samples$id, "quant.sf")
samples$files <- files

samples$names <- samples$id
se <- tximeta(samples)
se <- summarizeToGene(se)
ddsTxi <- DESeqDataSet(se, design = ~ condition)
```
# From count matrix (not run since we don't have matrix now)

```{r}
# cts <- as.matrix(read.csv(pasCts,sep="\t",row.names="gene_id"))
# coldata <- read.csv(pasAnno, row.names=1)
# coldata <- coldata[,c("condition","type")]
# coldata$condition <- factor(coldata$condition)
# coldata$type <- factor(coldata$type)
```

```{r}
dds <- DESeq(ddsTxi)


resultsNames(dds)
```
# Plot the estimates over the mean of scaled counts for each gene; Note that many of the plots in DESeq2 refer to “normalized counts”; here this just implies scaling the counts by the size factor, so that the differences affecting counts across samples are minimized.

```{r}
plotDispEsts(dds, ylim=c(1e-3, .5), xlim=c(5,1e5))
```

```{r}

res <- results(dds, name = "condition_NMD_OX_vs_HFD_OX")
```
+ MA plot of results

```{r}
plotMA(res, ylim=c(-8,8))
```

# Log fold change shrinkage for visualization and ranking

```{r}
resLFC <- lfcShrink(dds, coef="condition_NMD_OX_vs_HFD_OX", type="apeglm")
resLFC
```
+ MA plot Log fold change shrinkage for visualization and rank in

```{r}
plotMA(resLFC, ylim=c(-2,2))
```
# Plot counts

```{r}
plotCounts(dds, gene=which.min(res$padj), intgroup="condition")
```
# Cluster plot

```{r}
d <- plotCounts(dds, gene=which.min(res$padj), intgroup="condition", 
                returnData=TRUE)
library("ggplot2")
ggplot(d, aes(x=condition, y=count)) + 
  geom_point(position=position_jitter(w=0.1,h=0)) + 
  scale_y_log10(breaks=c(25,100,400))
```
# Heatmap

```{r}
library("pheatmap")
# this gives log2(n + 1)
ntd <- normTransform(dds)
# library("vsn")
# meanSdPlot(assay(ntd))
select <- order(rowMeans(counts(dds,normalized=TRUE)),
                decreasing=TRUE)[1:20]
df <- as.data.frame(colData(dds)[,c("condition", "sample_id")])
pheatmap(assay(ntd)[select,], cluster_rows=FALSE, show_rownames=FALSE,
         cluster_cols=FALSE, annotation_col=df)
```
# Tests of log2 fold change above or below a threshold


+ greaterAbs - $|\beta| > x$ - tests are two-tailed

+ lessAbs - $|\beta| < x$ - p values are the maximum of the upper and lower tests

+ greater $\beta > x$
+ less - $\beta < x$

```{r}
par(mfrow=c(2,2),mar=c(2,2,1,1))
ylim <- c(-2.5,2.5)
resGA <- results(dds, lfcThreshold=.5, altHypothesis="greaterAbs")
resLA <- results(dds, lfcThreshold=.5, altHypothesis="lessAbs")
resG <- results(dds, lfcThreshold=.5, altHypothesis="greater")
resL <- results(dds, lfcThreshold=.5, altHypothesis="less")
drawLines <- function() abline(h=c(-.5,.5),col="dodgerblue",lwd=2)
plotMA(resGA, ylim=ylim); drawLines()
plotMA(resLA, ylim=ylim); drawLines()
plotMA(resG, ylim=ylim); drawLines()
plotMA(resL, ylim=ylim); drawLines()
```

